---
title: "Cytoscape"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, message=FALSE, warning=FALSE, echo = FALSE)
```

```{r, eval = TRUE}
if (!requireNamespace("BiocManager", quietly = TRUE)){
    install.packages("BiocManager")
}

if (!requireNamespace("ComplexHeatmap", quietly = TRUE)){
  BiocManager::install("ComplexHeatmap")
}
library(ComplexHeatmap)

if (!requireNamespace("circlize", quietly = TRUE)){
  BiocManager::install("circlize")
}
library(circlize)

if (!requireNamespace("edgeR", quietly = TRUE)){
  BiocManager::install("edgeR")
}

library(edgeR)

if (!requireNamespace("knitr", quietly = TRUE)){
  BiocManager::install("knitr")
}

if (!requireNamespace("limma", quietly = TRUE)){
  BiocManager::install("limma")
}
library(limma)

if (!requireNamespace("RCy3", quietly = TRUE)){
  BiocManager::install("RCy3")
}
library(RCy3)

if (!requireNamespace("RCurl", quietly = TRUE)){
  BiocManager::install("RCurl")
}
library(RCurl)

if (!requireNamespace("Biobase", quietly = TRUE)){
  BiocManager::install("Biobase")
}
library(Biobase)

if (!requireNamespace("ggplot2", quietly = TRUE)){
  install.packages("ggplot2")
}
library(ggplot2)

if (!requireNamespace("httr", quietly = TRUE)){
  install.packages("httr")
}
library(httr)

if (!requireNamespace("RJSONIO", quietly = TRUE)){
  install.packages("RJSONIO")
}
library(RJSONIO)
```

Though I can run GSEA successfully in docker container, for reasons I don't know, when I tried to run docker by code in the docker container. It always returns

"Error in curl::curl_fetch_memory(url, handle = handle) : 

Failed to connect to localhost port 1234: Connection refused"

So I decided to create the html notebook for cytoscape part outside the container, then merge the resulstant html file directly into the final hrml report.


```{r checkcytoscape, eval = FALSE}
# Basic settings
port.number = 1234
base.url = paste("http://localhost:", toString(port.number), "/v1", sep="")

#print(base.url)

version.url = paste(base.url, "version", sep="/")
cytoscape.open = TRUE

tryCatch(expr = { GET(version.url)}, 
         error = function(e) { return (cytoscape.open = FALSE)}, finally =function(r){ return(cytoscape.open = TRUE)})

if(!cytoscape.open){
  #try and launch cytoscape
 print("Cytoscape is not open.  Please launch cytoscape.")
} else{
  cytoscape.version =  GET(version.url)
  cy.version = fromJSON(rawToChar(cytoscape.version$content))
  print(cy.version)
  
}

```

## Original EM before Manual Layout

```{r EMoriginal,eval = FALSE}
pvalue_gsea_threshold <- 0.005
qvalue_gsea_threshold <- 0.005

similarity_threshold <- "0.5"
similarity_metric = "OVERLAP"

cur_model_name <- "enrichment map"

#gsea_results_path <- file.path(gsea_result,"edb")
#gsea_results_filename <- file.path(gsea_results_path,"results.edb")

gmt_gsea_file <- file.path(dest_gmt_file)
#gsea_ranks_file <- file.path(gsea_results_path,list.files(gsea_results_path,pattern=".rnk"))

current_network_name <- paste(cur_model_name,pvalue_gsea_threshold,qvalue_gsea_threshold,sep="_")

em_command = paste('enrichmentmap build analysisType="gsea"',
                   'gmtFile=',gmt_gsea_file,
                   'pvalue=',pvalue_gsea_threshold, 
                   'qvalue=',qvalue_gsea_threshold,
                   'similaritycutoff=',similarity_threshold,
                   'coefficients=',similarity_metric,'ranksDataset1=E:/OneDrive/Document/UofT/LastYear/Winter/BCB420/submission/my_analysis.GseaPreranked.1590521512636/edb/rank.rnk', 
                   'enrichmentsDataset1=E:/OneDrive/Document/UofT/LastYear/Winter/BCB420/submission/my_analysis.GseaPreranked.1590521512636/edb/results.edb',
                   sep=" ")
response <- commandsGET(em_command)
```

```{r}
fitContent()

output_network_file <- file.path(getwd(),"OriginalEM.png")

if(file.exists(output_network_file)){
  #cytoscape hangs waiting for user response if file already exists.  Remove it first
  response <- file.remove(output_network_file)
} 

response <- exportImage(output_network_file, type = "png")

```
